What's hot and what's not:\\Windowed developer topic analysis

### 2009-06-25 lda-paper.tex
some minor changes

@@

\hlb{
Topic analysis extracts independant word distributions (topics) from
documents (commit log comments).
Ideally these extracted topics }
correlate with actual development topics that the developers discussed
during the development of the software project. 
Topic analysis often allows for a single document, such as a commit
\hlb{
message, to be related to multiple topics. Documents represented as a mixture of topics maps well to commits
}
\hla{
to source code, which often have more than one purpose.  A topic
represents both a word distribution and a group of commit log comments
that are related to each other by their content.  In this case a topic
is a set of tokens extracted from commit messages found within a
projects source control system (SCS). 
}

@@

\hla{
Imagine that we are analyzing one month of development in a project
where the main focus was on bug fixing, adding a new GUI and
documentation updates relating to the system and the GUI. To use LDA
we would get the word distributions of each commit messages, pass them to
LDA and tell LDA to find some topics, in this case we
\hlb{
 in this case we chose to concentrate on four topics.
} The four
topics it discovered would ideally be bug fixing, GUI, documentation
and other. These topics are essentially word distributions, for
example the bug fixing topic would include words like \emph{bug},
\emph{fix}, \emph{failure} and \emph{ticket}. The GUI topic would
include words like \emph{widget}, \emph{panel}, \emph{window}, and
\emph{menu}. The documentation topic would include words like
\emph{section}, \emph{figure}, and \emph{chapter}. The other topic
would include words found in topics generally independant of the other
topics. If we analyzed a commit of a bug fix it would probably contain
words like \emph{ticket} and thus heavily associate with the bug fix
topic. A commit that was about user documentation regarding the GUI
would be associated with the documentation topic and the GUI topic
because it shared words from both topics. Where as a topic dealing
neither bugs, GUIs or documentation might be associated with the other
topic. This example demonstrates how LDA can associate one document (a
commit message) with more than one topic, and how topics are related to
documents via their word distributions. LDA also would discover these
independant topics in an unsupervised manner. All the end user has to
do is name the topic based upon the most frequent words in that topic,
which is a word distribution.
}

@@

Once we have our data windowed, we apply our Topic Analysis tool to
each window and extract the top $N$ topics. We used $20$ topics in our
\hla{
case studies because we experimented before and that $20$ was about
all we can handle before too many topics seemed too indistinct from
each other to differeniate.

@@

Our input for topic similarity is the top 10 most common words of each
topic.  Each topic is compared to each other topic in the system,
given a certain threshold of top 10 word similarity, such as 8 out of
\hla{
top 10 matching words.  10 words were chosen because often people care
about top 10 comparison and it allowed for some common words to exist
but not cause topics to match so closely.  We then apply the
transitive closure on similar topics to this topic similarity matrix;
this is similar to modelling topics as nodes and similarity as arcs,
then fill flooding along the similarity arcs until we have partitioned
the topics into clusters of similar topics. Figure \ref{fig:closure}
illustrates clustering of topics by topic similarity. These clusters
of topics are called trends. 
Trends are interesting because they that topic during development has occured over more thanone period.
}

@@

\shrinkit
\SubSection{Visualization}
\shrinkit
%       1. [ ] need the visualizer
\hlb{ Visualization is an integral part of our topics analysis which
  allows us to quickly explore the topics and trends of a project.
  Visualization provides us with a framework that allows us to
  visually answer many questions about: the spread of topics, how many
  topics were independant per period, what the repeating trends were,
  if the trends dominated or did local topics dominate, how continuous
  trends were, and what the topic text in topic, trend, or period was.
  Since we have multiple questions to answer we have multiple
  visualization, described within this section, that help answer
  different questions.  }

@@

\hlb{
The \emph{trend-time-line} (Figure \ref{fig:trendtimeline})
attempts to show a summary of trends seperated from single topics. This view shows how a trend persists across time, which aids time-wise analysis of trends.
}
The \emph{trend-time-line} displays repeating trends more clearly by
dedicating a horizontal line for trend segments belonging to one
trend. Therefore if a trend contains discontinuous segments then
the segments appear on the same line.  However, the least common
trends need to be pruned away or the view will be very long. 
Thus the \emph{trend-time-line} view is used to analyze trends across time.

@@

When we examined PostgreSQL's history from 1996 to 2004, which includes
over $20 000$ commits. We did not find many trends with two or more
topics, using a similarity of $7/10$. $7/10$ was chosen because it
preserve indepedent topics but seemed to be a threshold
value where more serious trends started to appear.

Those trends that we did find
were not very large, lasting only 3 months at most. The first and second
largest trends directly referenced two external developers:
Dal Zotto, Dan McGuirk. The fifth largest trend related to work by
PostgreSQL developer D'Arcy. Other topics of the larger trends were
changes to the ``to do'' list, and time string formating topics
relating to timezones.

@@

Windowed topic analysis demonstrated its ability to hilight local topics
and indentify global trends. This was shown in our case study of MaxDB
7.500. Global topic analysis missed important topics such as
\emph{ODBC} while windowed topic analysis identified them.

>>>>

% \newcommand{\hl}[1]{#1}
% \newcommand{\hla}[1]{#1}
% \newcommand{\hlb}[1]{#1}
% \newcommand{\hlc}[1]{#1}

@@

\hlb{
Topic analysis extracts independent word distributions (topics) from
documents (commit log comments).
Ideally these extracted topics }
correlate with actual development topics that the developers discussed
during the development of the software project. 
Topic analysis often allows for a single document, such as a commit
\hlb{
message, to be related to multiple topics. Documents represented as a mixture of topics maps well to commits
}
\hla{
to source code, which often have more than one purpose.  A topic
represents both a word distribution and a group of commit log comments
that are related to each other by their content.  In this case a topic
is a set of tokens extracted from commit messages found within a
projects source control system (SCS). 
}

@@

\hla{
Imagine that we are analyzing one month of development in a project
where the main focus was on bug fixing, adding a new GUI and
documentation updates relating to the system and the GUI. To use LDA
we would get the word distributions of each commit messages, pass them to
LDA and tell LDA to find some topics, in this case we
\hlb{
 in this case we chose to concentrate on four topics.
} The four
topics it discovered would ideally be bug fixing, GUI, documentation
and other. These topics are essentially word distributions, for
example the bug fixing topic would include words like \emph{bug},
\emph{fix}, \emph{failure} and \emph{ticket}. The GUI topic would
include words like \emph{widget}, \emph{panel}, \emph{window}, and
\emph{menu}. The documentation topic would include words like
\emph{section}, \emph{figure}, and \emph{chapter}. The other topic
would include words found in topics generally independent of the other
topics. If we analyzed a commit of a bug fix it would probably contain
words like \emph{ticket} and thus heavily associate with the bug fix
topic. A commit that was about user documentation regarding the GUI
would be associated with the documentation topic and the GUI topic
because it shared words from both topics. Where as a topic dealing
neither bugs, GUIs or documentation might be associated with the other
topic. This example demonstrates how LDA can associate one document (a
commit message) with more than one topic, and how topics are related to
documents via their word distributions. LDA also would discover these
independent topics in an unsupervised manner. All the end user has to
do is name the topic based upon the most frequent words in that topic,
which is a word distribution.
}

@@

Once we have our data windowed, we apply our Topic Analysis tool to
each window and extract the top $N$ topics. We used $20$ topics in our
\hla{
case studies because we experimented before and that $20$ was about
all we can handle before too many topics seemed too indistinct from
each other to differentiate.

@@

Our input for topic similarity is the top 10 most common words of each
topic.  Each topic is compared to each other topic in the system,
given a certain threshold of top 10 word similarity, such as 8 out of
\hla{
top 10 matching words.  10 words were chosen because often people care
about top 10 comparison and it allowed for some common words to exist
but not cause topics to match so closely.  We then apply the
transitive closure on similar topics to this topic similarity matrix;
this is similar to modelling topics as nodes and similarity as arcs,
then fill flooding along the similarity arcs until we have partitioned
the topics into clusters of similar topics. Figure \ref{fig:closure}
illustrates clustering of topics by topic similarity. These clusters
of topics are called trends. 
Trends are interesting because they that topic during development has occurred over more than one period.
}

@@

\shrinkit
\SubSection{Visualization}
\shrinkit
%       1. [ ] need the visualizer
\hlb{ Visualization is an integral part of our topics analysis which
  allows us to quickly explore the topics and trends of a project.
  Visualization provides us with a framework that allows us to
  visually answer many questions about: the spread of topics, how many
  topics were independent per period, what the repeating trends were,
  if the trends dominated or did local topics dominate, how continuous
  trends were, and what the topic text in topic, trend, or period was.
  Since we have multiple questions to answer we have multiple
  visualization, described within this section, that help answer
  different questions.  }

@@

\hlb{
The \emph{trend-time-line} (Figure \ref{fig:trendtimeline})
attempts to show a summary of trends separated from single topics. This view shows how a trend persists across time, which aids time-wise analysis of trends.
}
The \emph{trend-time-line} displays repeating trends more clearly by
dedicating a horizontal line for trend segments belonging to one
trend. Therefore if a trend contains discontinuous segments then
the segments appear on the same line.  However, the least common
trends need to be pruned away or the view will be very long. 
Thus the \emph{trend-time-line} view is used to analyze trends across time.

@@

When we examined PostgreSQL's history from 1996 to 2004, which includes
over $20 000$ commits. We did not find many trends with two or more
topics, using a similarity of $7/10$. $7/10$ was chosen because it
preserve independent topics but seemed to be a threshold
value where more serious trends started to appear.

Those trends that we did find
were not very large, lasting only 3 months at most. The first and second
largest trends directly referenced two external developers:
Dal Zotto, Dan McGuirk. The fifth largest trend related to work by
PostgreSQL developer D'Arcy. Other topics of the larger trends were
changes to the ``to do'' list, and time string formating topics
relating to time-zones.

@@

Windowed topic analysis demonstrated its ability to hi-light local topics
and identify global trends. This was shown in our case study of MaxDB
7.500. Global topic analysis missed important topics such as
\emph{ODBC} while windowed topic analysis identified them.
